---
title: "Early Results"
author: "Tamara Dzolic"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Things to do for friday (Early Results): - descriptive/summary stats;
data visuals - statistical analysis: - Pearsons correaltion -
multicolinearity; V/F - Something you started \* Ideally lots of stats

#We know polarization has occured, So NOW WE LOOK AT HOW IT AFFECTS
WOMEN

#Load in libraries

```{r}
library(dplyr)
library(lattice)
library(ggplot2)
library(pacman)
library(tidyr)
library(vtable)
library(glmnet)
```

#Check the working directory

```{r}
getwd()
```

#Load in data

```{r}
#Political Polarization Data for each state
state_ideology <- read.csv("stateideology_v2018.csv")

#US Women Peace and Security Index Data
StatTable1 <- read.csv("StatTable1.csv")
```

#View both datasets
```{r}
View(StatTable1)
View(state_ideology)
colnames(state_ideology)
```
```{r}
View(StatTable1)
mean(StatTable1$US.WPS.Index.2020)
```


#Do a join on Georgetowns Womens Peace and Security Index Data
#And on the state legislatures data 

```{r}

state_ideology <- state_ideology %>%
  rename(State = ï..State)

#Outer join on both datasets
df_join = merge(x = state_ideology, 
                y = StatTable1, by = "State", all=TRUE)

#View results
View(df_join)

#Change column names 
df_join <- df_join %>%
  rename(WPS_Index_rank = ï..US.WPS.Index.rank,
         WPS_Index_2020 = US.WPS.Index.2020,
         Working_poor = Working.poor,
         State_Legislature = State.legislaturec,
         College_degree = College.degreec,
         Reropuctive_healthcare_access = Reproductive.healthcare.access,
         Legal_protection = Legal.protectiona,
         Discriminatory_norms = Discriminatory.normsb,
         Maternal_mortality = Maternal.mortalityc,
         Intimate_partner_violence = Intimate.partner.violence,
         Gun_deaths = Gun.deathsc.d,
         Healthcare_affordability = Healthcare.affordability,
         Community_safety = Community.safetyb
         )

#Check to make sure column names were changed 
colnames(df_join)
```
#Lets do some summary stats

```{r}
st(df_join)
summary(df_join)
```

```{r}
ggplot(data = state_ideology, aes(Government_Ideology)) + geom_histogram(aes(binwidth = 0.1), colour = "black", fill = "white") + geom_density(alpha = .2, fill="#FF6666") + xlab("Governemnt Ideologies") + ylab("Frequency")
```

```{r}
both_ideologies <- ggplot(data = state_ideology, aes(x = Num_code_for_state, y = Government_Ideology)) + geom_point()
  
both_ideologies
```

#Pearson correlation coefficient between citizens and state legislatures ideology

```{r}
cor(state_ideology$Citizen_Ideology, state_ideology$Government_Ideology, method = c("pearson"))
cor.test(state_ideology$Citizen_Ideology, state_ideology$Government_Ideology, method=c("pearson"))
```

#Moderate correlation score between citizens ideology and goverments.
#It matters who you vote for, but it matters more how the person you vote for votes

#Lets also do a pearsons correlation coefficient on Government Ideology scores and Georgetowns rank for women in the U.S

```{r}
cor(df_join$Government_Ideology, df_join$WPS_Index_2020, method = c("pearson"))
cor.test(df_join$Government_Ideology, df_join$WPS_Index_2020, method=c("pearson"))
```

#Lets do a LASSO Regression now to test the strengtth of this relationship

```{r}
colnames(df_join)
```

```{r MLR}
fit <- lm(Government_Ideology ~ Employment +
                             Working_poor +
                             State_Legislature +
                             College_degree +
                             Reropuctive_healthcare_access +
                             Legal_protection +
                             Discriminatory_norms +
                             Intimate_partner_violence +
                             Gun_deaths +
                             Healthcare_affordability +
                             Community_safety,
                             data = df_join)

summary(fit)
```



```{r Lasso}
#Independent variable (predictor variable)
#y <- data.matrix(df_join$Government_Ideology)

#Dependent Variable (matrix)
#x <- data.matrix(df_join[, c('Employment',
         #                    'Working_poor',
        #                     'State_Legislature', 
       #                      'College_degree',
      #                       'Reropuctive_healthcare_access',
     #                        'Legal_protection',
    #                         'Discriminatory_norms',
   #                          'Intimate_partner_violence',
  #                           'Gun_deaths',
 #                            'Healthcare_affordability',
#                             'Community_safety'
 #                            )])

#print(y)

#lambda <- 0.01

#best_model <- glmnet(x, y, alpha = 1, lambda = lambda)
#coef(best_model)
```

